                     HGMD Human Mutation Analysis
                      Kumar Lama, Dr. Ananda Mondal
---
title: "HGMD Human Mutation Data Analysis"
output: html_document
---
List of **Databases** in the Server:
-hgmdb_pro
-hgmd_phenbase
-hgmd_snp
-hgmd_views

Required Libraries 
'''
library(DBI)
library(RMySQL)
'''
Basic analysis for table Splice
```{r}
#install.packages("RMySQL",type = 'source')

#MySqL in R
#Install from source, package not available 
#Sys.setenv(PKG_CPPFLAGS = "-I/usr/local/include/mysql")
# Sys.setenv(PKG_LIBS = "-L/usr/local/lib -lmysqlclient")
suppressMessages(library(DBI))
suppressMessages(library(RMySQL))
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
#Connect to database on local host
dbcon.hgmd <- dbConnect(MySQL(),user = "root", password = "paparam", host = "localhost", db = "hgmd_pro")

tables <- dbListTables(dbcon.hgmd)

length(tables)
# code to display table name, number of fields and number of records
for (files in tables){
  tbl.splice <- dbReadTable(conn = dbcon.hgmd, name = files) #read Table called splice from db
  cat(paste(files,length(tbl.splice),length(tbl.splice[,1]),"\n",sep = "\t"))
  write.table(data.frame(files,length(tbl.splice),length(tbl.splice[,1])),file="tablesInfo.csv",append=T,row.names = F,col.names = F,sep = ",")
}

tbl.splice <- dbReadTable(conn = dbcon.hgmd, name = "mutation") #read Table called splice from db
write.csv(tbl.splice, file = "splice_tbl.csv") #save the talbe splice into a file splice_tbl.csv

tbl.gene <- table(tbl.splice$gene)
tbl.genelessthan <- tbl.gene[tbl.gene < 30 ]
barplot(tbl.genelessthan,xlab = paste("Splice genes: ",toString(as.integer(100 *length(tbl.genelessthan)/length(tbl.gene))),"% of total",sep = " "), ylab = "Frequency < 30")

gene.group <- tbl.splice%>%
  subset(!is.na(gene))%>%
  group_by(gene)%>%
  summarise(count = n())
  
tail(gene.group[order(gene.group$count),],20) #top six high frequncy genes

gene.group$gene <- 1:1978 #number the genes for plotting
p <- ggplot(aes(x = gene, y = count), data = subset(gene.group,count > 100 )) +
  geom_point(aes(alpha = 1/25))+
  geom_smooth()+ xlab("gene index")+
  ylab("frequency of the splice gene < 30")

quantile(gene.group$count,probs = c(0,0.5,0.75,0.90,0.95,0.97,0.98,1))
summary(gene.group$count)

  ggplot(aes(x = gene, y = count), data= gene.group ) + 
  geom_point()
```
**Descriptive Analysis for Mutation**
```{r}
tbl.mutation <- dbReadTable(conn = dbcon.hgmd, name = "mutation")#read mutation from db
write.csv(tbl.mutation, file = "mutation.csv") #save the talbe splice into a file splice_tbl.csv

tbl.mutationgene <- table(tbl.mutation$gene)
tbl.genelessthan <- tbl.gene[tbl.gene > 30 ]
barplot(tbl.genelessthan,xlab = paste("All Mutation genes, Total = ",sep = " ", toString(length(tbl.mutationgene))), ylab = "Frequency")

gene.group <- tbl.mutation%>%
  subset(!is.na(gene))%>%
  group_by(gene)%>%
  summarise(count = n())
  
tail(gene.group[order(gene.group$count),])
gene.group$gene <- 1:5559 #number the genes for plotting
p <- ggplot(aes(x = gene, y = count), data = subset(gene.group,count < 30)) +
  geom_point(aes(alpha = 1/25))+
  geom_smooth()+ xlab("gene index")+
  ylab("frequency of the mutation gene < 30")

quantile(gene.group$count,probs = c(0, 0.5,0.75,0.90,0.95,0.97,0.98,1))
summary(gene.group$count)



tbl.amino <- dbReadTable(conn = dbcon.hgmd, name = "amino")#read amino from db
tbl_genes <- table(tbl.splice[,"gene"])
names(tbl_genes) #table of the genes
length(tbl_genes) #total gene counts
length(tbl.splice[1,]);length(tbl.splice[,1])
names(tbl.splice)
set.seed(5555)
#splice.con[sample(14849,100),c("gene","type","base")]
#splice.columns <- dbGetQuery(conn = dbcon.hgmd, "show tables from hgmd_pro") 
dbDisconnect(dbcon.hgmd)#Disconnect the database

```
Analysis of Disease and genes
First, table is Grouped by disease and genes 

```{r}

gene.group <- tbl.splice%>%
  subset(!is.na( gene))%>%
  group_by(disease,gene)%>%
  summarise(count = n())%>%
  ungroup()
unique.diseases <- unique(gene.group$disease)
write.csv(gene.group,file="splice-disease-gene.csv")
gene.group$disease <- 1:3334
ggplot(data = subset(gene.group, count < 20), aes(x = disease, y = count)) + geom_point(aes(color = gene))
```

```{r}
tbl =  dbReadTable(conn = dbcon.hgmd, name = "goterms")#read goterms table from db
#group the table by go-term-acc, and count each terms frequency
gotermsCount<- tbl%>% # passes tbl to subset
  subset(!is.na(go_term_acc))%>% # passes non na go-terms to group_by function
  group_by(go_term_acc)%>% # group the data frame by go-term-acc and pass to summarise function
  summarise(frequency = n())%>% # summarise by counting frequency
  arrange(frequency) # order the new grouped table by descending frequency count 


goterms.group.type <- tbl%>%
  subset(!is.na(go_term_type))%>%
  group_by(go_term_acc,go_term_name)%>%
  summarise(frequency = n())%>%
  ungroup()%>%
  arrange(frequency)
  
write.table(goterms.group.type,file = "goterms-name.txt",sep = "\t")
write.table(gotermsCount,file="goterms-count.txt",sep = "\t")
with(tbl, table(go_term_type))
mean(gotermsCount$frequency)
```



